Model-based variable selection for regression with highly correlated variables.

### 2014-05-22 www/article13.tex
up
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/correg@620 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

	\subsection{The datasets}	
	Now we have defined the model and the way to obtain it, we can have a look on some numerical results to see if \textsc{CorReg} 	keeps its promises.
	The \textsc{CorReg} package has been tested on simulated datasets. 
Section \ref{compZ} shows the results obtained in terms of $\hat{S}$. Sections \ref{tableMSEsimtout} and \ref{tableMSEsimgauche} show the results obtained using only \textsc{CorReg}, or \textsc{CorReg} combined with other methods. Tables give both mean and standard deviation of the observed Mean Squared Errors (MSE) on a validation sample of $1 000$ individuals. For each simulation,  $p=40$, the $R^2$ of the main regression is $0.4$, variables in $\boldsymbol{X}_f$ follow Gaussian mixture models of $\lambda=5$ classes which means follow Poisson's law of parameter $\lambda=5$ and which standard deviation is $\lambda$. The $\beta_j$ and the coefficients of the $\boldsymbol{\alpha}_j$ are generated according to the same Poisson law but with a random sign. $\forall j \in I_r, p_1^j=2$ (sub-regressions of length 2) and we have $p_r=16$ sub-regressions. The datasets were then scaled so that covariates $X_r$ don't have a greater variance or mean.
	We used \textsc{Rmixmod} to estimate the densities of each covariate. For each configuration, the MCMC walk was launched on $10$ initial structures with a maximum of 1 000 steps each time.
	When $n<p$, a frequently used method is the Moore-Penrose generalized inverse \cite{katsikis2008fast}, thus OLS can obtain some results even with $n<p$. %(see tables \ref{YXlinOLS} and \ref{YX2linOLS} ).
	When using penalized estimators for selection, a last Ordinary Least Square step is added to improve estimation because penalisation is made to select variables but also shrinks remaining coefficients. This last step allows to keep the benefits of shrinkage (variable selection) without any impact on remaining coefficients (see \cite{SAM10088}) and is applied for both classical and marginal model.
	We compare different methods with and without CorReg as a pretreatment. All the results are provided by the CorReg package.
		\subsection{Finding the structure}
		\subsubsection{How to evaluate found structure?}
			The first criterion is $\psi(\boldsymbol{X},S)$ which is maximized in the MCMC. But in our case, it is estimated by the likelihood (see (\ref{approxBIC}))whose value don't have any intrinsic meaning. To show how far the found structure is from the true one in terms of $S$ we define some indicators to compare the true model $S$ and the found one $\hat{S}$.
			Global indicators :
			\begin{itemize}
				\item $TL$ (True left) : the number of found dependent variables that really are dependent $TL=|I_r\cap \hat{I}_r|$ 
				\item $WL$ (Wrong left) : the number of found dependent variables that are not dependent $WL=|\hat{I}_r|-TL$
				\item $ML$ (Missing left) : the number of really dependent variables not found $ML=|I_r|-TL$
				\item $\Delta p_r$ : the gap between the number of sub-regression in both model : $\Delta p_r=|I_r|-|\hat{I}_r|$. The sign defines if $\hat{S}$ is too complex or too simple
				\item $\Delta compl$ : the difference in complexity between both model : $\Delta compl=\sum_{j \in p_r}p_f^j-\sum_{j \in \hat{p}_r}\hat{p}_f^j$
			\end{itemize}
		\subsubsection{Results on $S$}	\label{compZ}
%In table \ref{compZvrai} 
We compare found structures in different contexts with both Uniform and Hierarchical Uniform {\it a priori} law on $P(S)$, noted respectively $BIC$ and $BIC_+$. 
We see that usage of $BIC_+$ gives sparser models even with $\max (\hat{p}_f^j)=5$. Moreover, this constraint is not active because observed complexities are smaller (this constraint only serves to accelerate the walk by reducing the dimension of $\mathcal{S}$ because each configuration was computed a hundred times), meaning that the stronger penalty implied by $BIC_+$ really is efficient. The datasets used for $BIC$ and $BIC_+$ are the same to keep the comparison  meaningful. %These datasets and $\hat{S}$ are used for   %tables \ref{YXlinOLS} to \ref{YX2linstep}.

 It is also notable that $BIC_+$ has a greater computational cost than $BIC$. 
	We notice that the MCMC is faster when there are numerous correlations (rejecting more candidates). 
\begin{table}[h!]
\centering
\begin{tabular}{|c|c|c|c|c|c|c|c|c|c|}
\hline
\multicolumn{3}{|c}{Configuration}  &  \multicolumn{2}{|c}{Computing Time}  & \multicolumn{3}{|c}{Quality} & \multicolumn{2}{|c|}{Complexity}\\
\hline
$n$ & $p_r$ & $P(S)$ &  Time Mixmod  & Time MCMC  & $TL$ & $WL$ & $ML$ & $\Delta p_r$ & $\Delta compl$ \\
\hline %global30p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_51_44.csv
30 & 0 & $BIC$ &0.4104 & 3.2428 & 0 & 5.43 & 0 & -5.43 & 22.55  \\
& & & (0.0275) & (0.3711) & (0) & (1.9346) & (0) & (1.9346) & (8.0884) \\
 &  & $BIC_+$ &0.4104 & 8.2338 & 0 & 0.53 & 0 & -0.53 & 2.27  \\
& & & (0.0275) & (0.9045) & (0) & (0.7844) & (0) & (0.7844) & (3.3024) \\
\hline %global30p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_49_58.csv
30 & 16 & $BIC$ &0.4182 & 2.7735 & 10.96 & 5.93 & 4.98 & -0.95 & 38.16  \\
& & & (0.0329) & (0.1498) & (1.9844) & (2.0313) & (1.9948) & (0.9987) & (6.499) \\
 &  & $BIC_+$ &0.4182 & 4.1876 & 11.61 & 4.57 & 4.33 & -0.24 & 16.48  \\
& & & (0.0329) & (0.2178) & (1.8743) & (1.9502) & (1.8752) & (0.4948) & (5.4892) \\
\hline %global30p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_19_50_41.csv
30 & 32 & $BIC$ & 0.4456 & 2.9154 & 25.23 & 1.92 & 6.5 & 4.58 & 28  \\
& & & (0.0429) & (0.1331) & (1.4761) & (1.0888) & (1.4668) & (0.9866) & (5.0831) \\
 &  & $BIC_+$ & 0.4456 & 4.0233 & 16.96 & 3.04 & 14.77 & 11.73 & 4.35  \\
& & & (0.0429) & (0.1091) & (1.3993) & (1.3993) & (1.4692) & (0.5478) & (5.8833) \\
\hline
\hline %global50p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_02_25_40.csv
50 & 0 & $BIC$ &0.5229 & 4.7068 & 0 & 4.2 & 0 & -4.2 & 13.35  \\
& & & (0.0519) & (0.5865) & (0) & (1.7233) & (0) & (1.7233) & (5.6468) \\
 &  & $BIC_+$ &0.5229 & 10.1198 & 0 & 0.13 & 0 & -0.13 & 0.32  \\
& & & (0.0519) & (0.5541) & (0) & (0.3667) & (0) & (0.3667) & (0.9732) \\
\hline %global50p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_50_48.csv
50 & 16 & $BIC$ &0.5205 & 3.3681 & 11.15 & 5.42 & 4.72 & -0.7 & 22.85  \\	
& & & (0.0451) & (0.3123) & (1.93) & (1.9132) & (1.886) & (0.7317) & (5.7742) \\
 &  & $BIC_+$ &0.5205 & 4.909 & 11.42 & 4.59 & 4.45 & -0.14 & 7.55  \\
& & & (0.0451) & (0.4556) & (1.9079) & (1.8968) & (1.8333) & (0.3487) & (4.0611) \\
\hline %global50p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_12mars2014_23_49_39.csv
50 & 32 & $BIC$ & 0.5833 & 3.2683 & 28.17 & 1.38 & 3.7 & 2.32 & 12.74  \\
& & & (0.0628) & (0.316) & (1.3711) & (0.9077) & (1.3143) & (0.8394) & (4.3359) \\
 &  & $BIC_+$ &0.5833 & 4.3599 & 17.27 & 2.73 & 14.6 & 11.87 & -2.61  \\
& & & (0.0628) & (0.3729) & (1.1708) & (1.1708) & (1.2792) & (0.338) & (4.4854) \\
\hline
\hline %global100p40ratio0BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_10_47_32.csv
100 & 0 & $BIC$ & 0.9623 & 12.9373 & 0 & 2.83 & 0 & -2.83 & 6.23  \\
& & & (0.077) & (1.7778) & (0) & (1.2953) & (0) & (1.2953) & (3.1999) \\
 &  & $BIC_+$ & 0.9623 & 20.9817 & 0 & 0.01 & 0 & -0.01 & 0.02  \\
& & & (0.077) & (1.9421) & (0) & (0.1) & (0) & (0.1) & (0.2) \\
\hline %global100p40ratio0.4BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_33_40.csv
100 & 16 & $BIC$ & 1.1223 & 6.9647 & 11.67 & 4.8 & 4.25 & -0.55 & 12.58  \\
& & & (0.1122) & (0.5473) & (2.0003) & (2.0646) & (1.956) & (0.7833) & (3.9471) \\
 &  & $BIC_+$ & 1.1223 & 8.8486 & 12.04 & 3.95 & 3.88 & -0.07 & 3.75  \\
& & & (0.1122) & (0.7174) & (1.9223) & (1.9404) & (1.9137) & (0.2564) & (2.2625) \\
\hline % global100p40ratio0.8BICTRUEBICstarTRUEp1max5lambda5100Amaxp2max100sigma_sousreg0.001maxit9000maxtry20max_compl1scaleTRUEcandidates-1_13mars2014_04_10_09.csv
100 & 32 & $BIC$ & 1.4343 & 5.9626 & 30.14 & 0.84 & 1.61 & 0.77 & 6.96  \\
& & & (0.2528) & (0.3136) & (1.3928) & (0.8495) & (1.2941) & (0.7086) & (3.0975) \\
 &  & $BIC_+$ & 1.4343 & 7.3741 & 17.49 & 2.51 & 14.26 & 11.75 & -3.76  \\
& & & (0.2528) & (0.2748) & (1.1849) & (1.1849) & (1.2441) & (0.4794) & (4.4859) \\
\hline
\end{tabular} 
\caption{Results of the Markov chain: Mean observed and standard deviation (sd). } \label{compZvrai}
\end{table}

@@

\clearpage
\subsubsection{$\boldsymbol{Y}$ depends only on covariates in $\boldsymbol{X_f}$ (best case for us)}	 \label{tableMSEsimgauche}
\begin{figure}[h!]
	\begin{minipage}[l]{.48\linewidth}
			\includegraphics[height=180px,width=245px]{figures/res_article/OLS_X1_MSE_NB.png} 
			\caption{Comparison of the MSE between OLS and CorReg+OLS}
	\end{minipage} \
   \begin{minipage}[r]{.48\linewidth}
			\includegraphics[height=180px,width=240px]{figures/res_article/OLS_X1_compl_NB.png} 
			\caption{Comparison of the complexities between OLS and CorReg+OLS} 
   \end{minipage}
\end{figure}

@@

\clearpage
\subsection{Results on prediction}
	\subsubsection{$\boldsymbol{Y}$ depends on all variables in $\boldsymbol{X}$}	 	
We then try the method with a response depending on all covariates (\textsc{CorReg} reduces the dimension and can't give the true model if there is a structure). The datasets used here were those from table \ref{compZvrai}.

>>>>

	\subsection{The datasets}	
	Now we have defined the model and the way to obtain it, we can have a look on some numerical results to see if \textsc{CorReg} 	keeps its promises.
	The \textsc{CorReg} package has been tested on simulated datasets. 
Section \ref{compZ} shows the results obtained in terms of $\hat{S}$. Sections \ref{tableMSEsimtout} and \ref{tableMSEsimgauche} show the results obtained using only \textsc{CorReg}, or \textsc{CorReg} combined with other methods. Tables give both mean and standard deviation of the observed Mean Squared Errors (MSE) on a validation sample of $1 000$ individuals. For each simulation,  $p=40$, the $R^2$ of the main regression is $0.4$, variables in $\boldsymbol{X}_f$ follow Gaussian mixture models of $\lambda=5$ classes which means follow Poisson's law of parameter $\lambda=5$ and which standard deviation is $\lambda$. The $\beta_j$ and the coefficients of the $\boldsymbol{\alpha}_j$ are generated according to the same Poisson law but with a random sign. $\forall j \in I_r, p_1^j=2$ (sub-regressions of length 2) and we have $p_r=16$ sub-regressions. The datasets were then scaled so that covariates $X_r$ don't have a greater variance or mean.
	We used \textsc{Rmixmod} to estimate the densities of each covariate. For each configuration, the MCMC walk was launched on $10$ initial structures with a maximum of 1 000 steps each time.
	When $n<p$, a frequently used method is the Moore-Penrose generalized inverse \cite{katsikis2008fast}, thus OLS can obtain some results even with $n<p$. %(see tables \ref{YXlinOLS} and \ref{YX2linOLS} ).
	When using penalized estimators for selection, a last Ordinary Least Square step is added to improve estimation because penalisation is made to select variables but also shrinks remaining coefficients. This last step allows to keep the benefits of shrinkage (variable selection) without any impact on remaining coefficients (see \cite{SAM10088}) and is applied for both classical and marginal model.
	We compare different methods with and without CorReg as a pretreatment. All the results are provided by the CorReg package.
		\subsection{Finding the structure}
		\subsubsection{How to evaluate found structure?}
			The first criterion is $\psi(\boldsymbol{X},S)$ which is maximized in the MCMC. But in our case, it is estimated by the likelihood (see (\ref{approxBIC}))whose value don't have any intrinsic meaning. To show how far the found structure is from the true one in terms of $S$ we define some indicators to compare the true model $S$ and the found one $\hat{S}$.
			Global indicators :
			\begin{itemize}
				\item $TL$ (True left) : the number of found dependent variables that really are dependent $TL=|I_r\cap \hat{I}_r|$ 
				\item $WL$ (Wrong left) : the number of found dependent variables that are not dependent $WL=|\hat{I}_r|-TL$
				\item $ML$ (Missing left) : the number of really dependent variables not found $ML=|I_r|-TL$
				\item $\Delta p_r$ : the gap between the number of sub-regression in both model : $\Delta p_r=|I_r|-|\hat{I}_r|$. The sign defines if $\hat{S}$ is too complex or too simple
				\item $\Delta compl$ : the difference in complexity between both model : $\Delta compl=\sum_{j \in p_r}p_f^j-\sum_{j \in \hat{p}_r}\hat{p}_f^j$
			\end{itemize}
		\subsubsection{Results on $S$}	\label{compZ}

@@

\clearpage
\subsection{Results on prediction}

\subsubsection{$\boldsymbol{Y}$ depends only on covariates in $\boldsymbol{X_f}$ (best case for us)}	 \label{tableMSEsimgauche}
\begin{figure}[h!]
	\begin{minipage}[l]{.48\linewidth}
			\includegraphics[height=180px,width=245px]{figures/res_article/OLS_X1_MSE_NB.png} 
			\caption{Comparison of the MSE between OLS and CorReg+OLS}
	\end{minipage} \
   \begin{minipage}[r]{.48\linewidth}
			\includegraphics[height=180px,width=240px]{figures/res_article/OLS_X1_compl_NB.png} 
			\caption{Comparison of the complexities between OLS and CorReg+OLS} 
   \end{minipage}
\end{figure}

@@

\clearpage
	\subsubsection{$\boldsymbol{Y}$ depends on all variables in $\boldsymbol{X}$}	 	
We then try the method with a response depending on all covariates (\textsc{CorReg} reduces the dimension and can't give the true model if there is a structure). The datasets used here were those from table \ref{compZvrai}.
